---
title: "Hospitalisations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hospitalisations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval=FALSE}
library(RSVsim)
```

We do not explicitly include hospitalisation in the transmission model, but can estimate the hospitalisations from the simulated incidence. Note the time until hospitalisation is therefore ignored.

## Probability of hospitalisation given infection

Hospitalisation rates (among the whole population) might be sampled but the total incidence not known. If so, we can use `RSVsim` to estimate the incidence in the setting in which the age specific hospitalisation rates were estimated and use this incidence and the hospitalisation rates to calculate the probability of hospitalisation given infection (infection hospitalisation rate; IHR). The probability of hospitalisation given unvaccinated infection for age group $i$ is calculated as $p(hosp|inf,i) = \frac{H_{i}}{I_{i}}$, where $H_{i}$ is the number of hospitalisations in age group $i$ and $I_{i}$ is the corresponding incidence. The age-specific hospitalisation rates per person is $h_{i}=\frac{H_{i}}{N_{i}}$, where $N_{i}$ is the population size, meaning $p(hosp|inf) = \frac{h_{i}}{I_{i}/N_{i}}$. Similarly, we can calculate $p(hosp|inf)$ in unvaccinated people assuming a given vaccine efficacy against hospitalisation conditional on infection for each vaccinated state ($VE^{C}_{v}$) and hospitalisation rates, thus,

$$
p(hosp|inf,i) = \frac{h_{i}}{\frac{I_{1i}}{N_{1i}} + \sum_{2}^{V} \frac{I_{vi}}{N_{vi}}(1 - VE^{C}_{vi})}
$$

With an overall vaccine efficacy against hospitalisation ($VE_{H}$) and vaccine efficacy against infection ($VE_{I}$) we can calculate the vaccine efficacy against hospitalisation conditional on infection thus,

$$
VE_{i}^{C} = 1 - \frac{1 - VE_{i}^{H}}{1 - VE_{i}^{I}}
$$
We provide the `RSVsim_VE_hosp_given_inf` function to calculate this; $VE^{I}$ should have the same ages $VE^{H}$. In this example we assume no existing vaccination. The default vaccination states are 1 (unvaccinated) and 2 (vaccinated). First, we use `RSVsim_run_model` to approximate the incidence.

```{r, eval=FALSE}

times <- seq(0, 365.25*11, 0.25)
warm_up <- 365.25 * 10
cohort_step_size <- 1/12 * 365.25

age_limits <- c(seq(0, 5, 1/12), seq(10, 75, 5))

contact_population_list <- RSVsim_contact_matrix(age_limits = age_limits)

parameters <- RSVsim_parameters(contact_population_list = contact_population_list)

sim_status_quo <- RSVsim_run_model(parameters = parameters,
                                   times = times,
                                   cohort_step_size = cohort_step_size,
                                   warm_up = warm_up)

```

We can then use the `RSVsim_prob_hosp_given_inf` function to estimate $p(hosp|inf)$ for given values of $h_{i}$ (note these must be per person). We must also provide the times over which the hospitalisations are estimated and the vaccine efficacy against hospitalisation for each vaccinated state.

```{r, eval=FALSE}

# age specific number of hospitalisations per person

hosp_rate <- c(5338.91375, 1751.145228, 715.9402014, 208.4621667, 10.86702733, 2.58129891, 2.321354938, 1.868628613,
                 1.320440418, 1.834551244, 1.883359955, 2.861263391, 4.744902927, 6.645799647, 9.348531561, 14.3538506,
                 22.62406397, 32.94296492, 87.63366855) / 100000

age_limits_hosp_rate <- c(0, 0.5, 1, 2, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70, 75) # lower age limits of the hospitalisation rate data

# vaccine efficacy against hospitalisation conditional on infection
VE_hosp <- matrix(rep(0.8, length(age_limits)), ncol = 1) # age-specific vaccine efficacy against hospitalisation
VE_inf <- parameters$VE_inf
VE_hosp_given_inf <- RSVsim_VE_hosp_given_inf(VE_hosp = VE_hosp, VE_inf = VE_inf)

prob_hosp_given_inf_list <- RSVsim_prob_hosp_given_inf(hosp_rate = hosp_rate,
                                                       age_limits_hosp_rate = age_limits_hosp_rate,
                                                       hosp_rate_min_time = 0, # time frame over which to sum incidence
                                                       hosp_rate_max_time = 365.25,
                                                       sim_status_quo = sim_status_quo,
                                                       VE_hosp_given_inf = VE_hosp_given_inf,
                                                       VE_hosp_vacc_states = c(2) # vector identifying the vaccination states that are vaccinated (position corresponds to the column position in VE_hosp)
                                                       )
```

## Hospitalisations

We also provide a function to estimate the number of hospitalisations in different scenarios from the `RSVsim_run_model` outputs and known $p(hosp|inf)$. First, we simulate a novel scenario and then use the `RSVsim_hospitalisations` function to estimate the number of hospitalisations over each time-step. In this example, we will use the $p(hosp|inf)$ values estimated using the `RSVsim_prob_hosp_given_inf` function. The `RSVsim_prob_hosp_given_inf` function returns a list of $p(hosp|inf,i)$ and the `sim_status_quo` model outputs with hospitalisations.

```{r, eval=FALSE}

#################################
##### transmission dynamics #####
#################################

# single 30 day vaccination campaign
vaccine_times <- c(0, c(90, 120) + 365.25 * 10)
vaccine_period <- diff(c(vaccine_times, max(vaccine_times)))
vaccine_period[vaccine_period == 0] <- 30
nVaccTimes <- length(vaccine_times)

nAges <- length(contact_population_list$age_limits)

# vaccination campaign target coverage: 80% in the oldest age group
vaccine_cov <- rbind(matrix(rep(0, (nAges-1) * nVaccTimes), nrow = (nAges-1), ncol = nVaccTimes),
                     c(0, 0.8, 0))

nVaccStates <- 2
VE_inf <- matrix(rep(0.5, nAges * (nVaccStates-1)), nrow = nAges) # vaccine efficacy against infection
gamma_vaccine <- 1 / (365.25 * 2)

vac_parameters <- RSVsim_parameters(overrides = list("vaccine_times" = vaccine_times,
                                                     "vaccine_period" = vaccine_period,
                                                     "nVaccTimes" = nVaccTimes,
                                                     "vaccine_cov" = vaccine_cov,
                                                     "nVaccStates" = nVaccStates,
                                                     "VE_inf" = VE_inf,
                                                     "gamma_vaccine" = gamma_vaccine),
                                    contact_population_list = contact_population_list)

sim_vaccination <- RSVsim_run_model(parameters = vac_parameters,
                                    times = times,
                                    cohort_step_size = cohort_step_size, # time at which to age people
                                    warm_up = warm_up)

############################
##### hospitalisations #####
############################

sim_vaccination <- RSVsim_hospitalisations(prob_hosp_given_inf = prob_hosp_given_inf_list[["prob_hosp_given_inf_uv"]],
                                           age_limits_hosp_rate = age_limits_hosp_rate,
                                           sim_scenario = sim_vaccination,
                                           VE_hosp_given_inf = VE_hosp_given_inf,
                                           VE_hosp_vacc_states = c(2))

```

