---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  autodep = TRUE
)
```

In this vignette, we give an overview of the mathematical model of RSV transmission and how to use this package to run the model.

```{r setup, eval=FALSE}
library(RSVsim)

```

## Contact matrix
 The [R/RSVsim_contact_matrix.R](https://github.com/IsaacStopard/RSVsim/blob/master/R/RSVsim_contact_matrix.R) function obtains contact matrices with age groupings that are less than one year. To do so, the socialmixr (<https://epiforecasts.io/socialmixr/>) package is used to obtain the contact matrix for the age groupings to the nearest whole year, using the POLYMOD data (<https://doi.org/10.1371/journal.pmed.0050074>). This contact matrix is calculated such that the total contacts in the population are symmetric (assuming population data from World Population Prospects of the United Nations). Within age groups less than one year, these contacts are divided evenly according to the size of the age group (in years). The function takes two arguments: (1) the country to obtain the POLYMOD contact data from and (2) the lower age limits of the age groups.

```{r contact_matrix, dependson="setup", eval=FALSE}
# specifying the some age limits to use with the model
age_limits <- c(seq(0, 2, 0.2), seq(10, 60, 20))

contact_population_list <- RSVsim_contact_matrix(country = "United Kingdom", age_limits = age_limits)

```

## Default parameters
To obtain a list of the default parameters to run the model use the [R/RSVsim_parameters.R](https://github.com/IsaacStopard/RSVsim/blob/master/R/RSVsim_parameters.R) function. The parameters can be changed by using the overrides argument; for example we override $b0$ (the transmission rate coefficient) with a value of $0.15$. The "fitted" argument refers to parameters that will be excluded - the function will not return any parameters named in this vector. Initial conditions are defined in the parameter list. If no initial conditions are given, it is assumed there is 0.1% prevalence in the primary infection compartment across all ages and everyone else is susceptible to primary infection. All new-borns are assumed to be in the susceptible to primary infection compartment.

```{r parameters, dependson="contact_matrix", eval=FALSE}
parameters <- RSVsim_parameters(overrides = list("b0" = 0.15, "b1" = 1, "phi" = 0), 
                                contact_population_list = contact_population_list)
```

## Running the model
To run the model the [R/RSVsim_run_model.R](https://github.com/IsaacStopard/RSVsim/blob/master/R/run_model.R) function is used. Cohort aging is used to age the population. At each time, a proportion of each cohort is aged, with this proportion determined by the length of time before people are aged (cohort_step_size) as a proportion of the total length of time in each age group.

```{r model_run, dependson="parameters", eval=FALSE}

times <- seq(0, 365 * 10, 0.25)

out <- RSVsim_run_model(parameters = parameters,
                        times = times, # maximum time to run the model for
                        cohort_step_size = 1/12 * 365, # time at which to age people\
                        warm_up = NULL)

```

The simulated number of people that are susceptible and infected, and whether the infection is primary or secondary is tracked in the model outputs. The incidence and detected incidence since the previous time-step are calculated and provided in the model outputs. To calculate the detected incidence age-specific detection probabilities are used, and in the default parameters we assume very few people (1%) over the age of 2 years old are tested. The age-specific prevalence is also calculated. 

The incidence and doses are given as the number that occurred for each time step. The time-steps when cohort ageing occurs are included in the simulation. Usually the difference between these and the time-steps provided in `times` are different meaning the incidence or doses are not consistently calculated over the time frame and can therefore appear to rapidly change. To estimate the incidence or doses over consistent time-steps we must therefore sum the incidence grouped by the primary simulation times.

```{r, dependson="model_run", eval=FALSE}

# times are incremented by 0.25 when running the model in this example

out |> dplyr::mutate(time_round = round(time / 0.25) * 0.25) |> 
  dplyr::group_by(time_round, vacc_state, age, age_chr) |> 
  dplyr::summarise(Incidence = sum(Incidence), doses = sum(doses))

```
